### 2022-09-11 main.tex
Create cover.tex

\documentclass[a4paper, 12pt]{article}
\usepackage[top=2.5cm, bottom=3cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[]{indentfirst}
\usepackage{graphicx}
\usepackage{subfiles}
\usepackage[brazilian]{babel}

\graphicspath{ {./images/} }

\setlength{\parskip}{0.5\baselineskip}
% \setlength{\parindent}{0pt}

\begin{document}

    \subfile{sections/cover.tex}

    \section{INTRODUÇÃO}
    \subfile{sections/introduction.tex}

    \section{METODOLOGIA} 
    \subfile{sections/methodology.tex}

    \section{RESULTADOS E DISCUSSÃO}
    \subfile{sections/results.tex}

    \section{CONCLUSÃO}
    \subfile{sections/conclusion.tex}

    \bibliographystyle{plain}
    \bibliography{refs.bib}

\end{document}